---
title: "Haunted Places in the United States"
subtitle: "INFO 526 - Fall 2024 - Project 01"
author: 
  - name: "The Data Drafters"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "The Haunted Places dataset is obtained from the TidyTuesday project, which collects public datasets for analysis. This particular dataset, released in October 2023, includes information on over 10,000 haunted locations across the United States.

Provenance: This dataset is crowdsourced from various public sources, including websites and historical records about haunted locations. The data is openly available for public use and analysis, making it suitable for this project.

Dimensions: The dataset contains 10,992 rows and 10 columns, with each row representing a unique haunted site across various U.S. states."
format:
   html:
    code-tools: true
    code-overflow: wrap
    embed-resources: true
editor: visual
execute:
  warning: false
  echo: false
---

## Abstract

The Haunted Places Dataset is a comprehensive compilation of 10,992 haunted locations across the United States, designed to facilitate exploration and analysis of cultural, historical, and geographical phenomena associated with hauntings. Each entry in the dataset includes key attributes such as city, state, description, and geographical coordinates (longitude and latitude), enabling a multi-dimensional analysis of haunted places. The dataset provides insights into regional distributions of hauntings, allowing for geographical mapping and the identification of hotspots. Additionally, it supports investigations into the historical significance of haunted locations by correlating narratives with geographical data. By leveraging this dataset, researchers and enthusiasts can delve into the rich tapestry of ghost stories, folklore, and historical events tied to these eerie sites, making it a valuable resource for academic studies, tourism development, and cultural analysis.

## Introduction

The Haunted Places Dataset provides a rich collection of haunted locations across the United States, with 10,992 rows and 10 columns detailing various attributes of these sites. Each entry represents a haunted location, offering insights through attributes such as the city, state, description, and geographical coordinates (longitude and latitude). This dataset is valuable for exploring the cultural, historical, and geographical aspects of hauntings, enabling analysis of trends and patterns in haunted locations.

The dataset offers opportunities for visualizing the geographical distribution of haunted places, understanding regional trends, and even linking hauntings to historical events or local folklore. Whether for research, storytelling, or tourism-related projects, this dataset provides a unique perspective on the haunted history of the United States.

```{r}
# Load necessary libraries

library(tidyverse)
library(ggplot2)
library(tidytext)
library(ggridges)
library(wordcloud2)
library(sf)  # For spatial analysis and maps
library(tmap)
library(leaflet)
library(ggrepel)
library(dplyr)
library(maps)
library(gganimate)
library(ggimage)
```

```{r}
# Load the dataset
haunted_places <- read_csv('haunted_places.csv')

```

```{r}
# View basic summary
summary(haunted_places)

```

```{r}
# Handling missing data: Drop rows with missing longitude or latitude
haunted_clean <- haunted_places %>%
  filter(!is.na(longitude) & !is.na(latitude))

```

## Dataset Description

The Haunted Places dataset is obtained from the [TidyTuesday project](https://github.com/rfordatascience/tidytuesday), which collects public datasets for analysis. This particular dataset, released in October 2023, includes information on over 10,000 haunted locations across the United States.

Provenance: This dataset is crowdsourced from various public sources, including websites and historical records about haunted locations. The data is openly available for public use and analysis, making it suitable for this project.

Dimensions: The dataset contains **10,992 rows** and **10 columns**, with each row representing a unique haunted site across various U.S. states.

## Question 1: Geographical Analysis

**Approach:**

To determine the most haunted states and cities, we first visualize the haunted places using longitude and latitude data. We also merge external population data to explore the relationship between state population and the number of haunted places. We plan to create heatmaps and scatter plots to showcase the density of haunted locations and their correlation with population figures.

## Visualizing haunted places across the U.S.

```{r}
# Define latitude and longitude limits for the contiguous USA
lat_limits <- c(24.396308, 49.384358)  # South and North bounds
long_limits <- c(-125.0, -66.93457)    # West and East bounds

# Remove outliers based on defined limits
haunted_clean_filtered <- haunted_clean %>%
  filter(latitude >= lat_limits[1], latitude <= lat_limits[2],
         longitude >= long_limits[1], longitude <= long_limits[2])

# Create a map outline of the USA
usa_map <- map_data("state")

# Plotting the outline map with haunted places
ggplot() +
  geom_polygon(data = usa_map, aes(x = long, y = lat, group = group), fill = "white", color = "black", size = 0.5) +  # Outline of the USA
  geom_point(data = haunted_clean_filtered, aes(x = longitude, y = latitude), 
             color = "darkred", size = 3, alpha = 0.7, shape = 21, fill = "red") +  # Points for haunted places
  labs(title = "Haunted Places in the USA", 
       subtitle = "Visualizing the Spooky Locations Across the States",
       x = "Longitude", 
       y = "Latitude") +
  theme_minimal(base_size = 14) +
  theme(panel.grid = element_blank(),
        plot.title = element_text(size = 20, face = "bold"),
        plot.subtitle = element_text(size = 14, face = "italic"),
        axis.text = element_text(size = 12))
```

## Haunted locations by state

```{r}
 # URL or local path for a ghost icon image
ghost_icon <- "https://img.icons8.com/?size=80&id=3T4nzAVK6IRW&format=png"  

# Top 10 haunted states
top_haunted_states <- haunted_clean %>%
  count(state_abbrev, sort = TRUE) %>%
  top_n(10, n) %>%
  mutate(type = "State")

# Top 10 haunted cities
top_haunted_cities <- haunted_clean %>%
  count(city, sort = TRUE) %>%
  top_n(10, n) %>%
  mutate(type = "City")

# Combine both dataframes
haunted_combined <- bind_rows(
  top_haunted_states %>% rename(location = state_abbrev),
  top_haunted_cities %>% rename(location = city)
)

# Add ghost icon URL to each row
haunted_combined <- haunted_combined %>%
  mutate(icon = ghost_icon)

# Lollipop plot with ghost icons
ggplot(haunted_combined, aes(x = reorder(location, n), y = n)) +
  geom_segment(aes(xend = location, y = 0, yend = n), color = "red", size = 1.5) +  
  ggimage::geom_image(aes(image = icon), size = 0.07) +  # Ghost icon instead of circles
  facet_wrap(~type, scales = "free", ncol = 2, strip.position = "top") +
  coord_flip() +
  labs(title = "Top 10 Haunted States and Cities by Number of Locations") + 
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold", size = 12),    # Bold labels for State and City
    axis.text.x = element_text(size = 10),                  # X-axis text size
    axis.text.y = element_text(face = "bold", size = 10),   # Bold Y-axis labels (State/City names)
    plot.title = element_text(size = 16, face = "bold"),    # Bold title
    legend.position = "none",                               # Remove the legend
    panel.spacing = unit(4, "lines"),                       # Gap between the two plots
    axis.title.x = element_blank(),                         # Remove x-axis title
    axis.title.y = element_blank()                          # Remove y-axis title
  )
```

## Correlating haunted places with state population

```{r}
# Load necessary libraries
library(dplyr)
library(readr)
library(janitor)

# Load population data
population_data <- read_csv("Haunted_Data_population.csv")  
# Clean column names in population data
population_data <- population_data %>%
  clean_names()

# Check column names to verify the state and population columns
print(colnames(population_data))

# Count the number of haunted places by state
haunted_counts_state <- haunted_places %>%
  group_by(state) %>%
  summarise(haunted_places_count = n()) %>%
  arrange(desc(haunted_places_count)) %>%
  head(10)  # Get the top 10 haunted states

# Extract the top 10 haunted states
top_haunted_states <- haunted_counts_state$state

# Check for unique state names in population data to match with haunted states
print(unique(population_data$name)) 

# Filter the population data for the top 10 haunted states
population_top_haunted <- population_data %>%
  filter(name %in% top_haunted_states) 


population_top_haunted <- population_top_haunted %>%
  rename(population = p1_001n)  # Adjust 'population_column_name' accordingly

# Merge haunted places count with the population data
haunted_population <- haunted_counts_state %>%
  left_join(population_top_haunted, by = c("state" = "name"))

# Display the merged result to verify
print(haunted_population)

# Check the structure to verify correct column names and types
str(haunted_population)

# Convert population and haunted places count to numeric (if needed)
haunted_population$population <- as.numeric(haunted_population$population)
haunted_population$haunted_places_count <- as.numeric(haunted_population$haunted_places_count)

# Check for NA values or non-numeric entries
summary(haunted_population$population)
summary(haunted_population$haunted_places_count)

# Remove rows with NA values
haunted_population_clean <- haunted_population %>%
  filter(!is.na(population) & !is.na(haunted_places_count))

# Check the cleaned data
print(haunted_population_clean)

# Calculate and print correlation between population and number of haunted places
correlation <- cor(haunted_population_clean$population, haunted_population_clean$haunted_places_count)
print(correlation)
```

```{r}
library(scales)
# Enhanced Bar plot with more visible population bars
enhanced_bar_plot <- ggplot(haunted_population_clean, aes(x = reorder(state, haunted_places_count))) +
  # Haunted places bar
  geom_bar(aes(y = haunted_places_count), stat = "identity", fill = "darkblue", width = 0.2) +
  geom_text(aes(y = haunted_places_count, label = haunted_places_count), 
            vjust = -0.5, color = "white", size = 3) +
  
  # Population bar scaled down and placed beside haunted places
  geom_bar(aes(y = population / 100000), stat = "identity", fill = "orange", width = 0.3, 
           position = position_nudge(x = 0.4)) +  # Decreased divisor for better visibility
  
  # Labels and title
  labs(
    title = "Population and Haunted Places in Top 10 States",
    x = "State",
    y = "Count (Number of Haunted Places)",
    fill = "Legend"
  ) +
  
  # Secondary y-axis for population with adjusted limits
  scale_y_continuous(
    name = "Number of Haunted Places",
    sec.axis = sec_axis(~ . * 100000, name = "Population", 
                        labels = label_number(big.mark = ",", decimal.mark = ".", scale = 1))  
  ) +
  
  # Aesthetics and theme adjustments
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    legend.position = "bottom"  # Position for legend
  ) +
  scale_fill_manual(values = c("darkblue" = "darkblue", "orange" = "orange"), 
                    labels = c("Haunted Places", "Population"), 
                    name = "Legend")  # Add a legend

# Display the enhanced bar plot
print(enhanced_bar_plot)


# Calculate and print correlation between population and number of haunted places
correlation <- cor(haunted_population_clean$population, haunted_population_clean$haunted_places_count)
print(correlation)
```

## Observation:

Initial observations suggest that certain states, like California and Texas, have a higher concentration of haunted sites.

Upon merging population data, we will examine whether more populous states also report more hauntings or if there are regional factors contributing to these concentrations.
