---
title: "Haunted Places in the United States"
subtitle: "Proposal"
author: 
  - name: "The Data Drafters"
    affiliations:
      - name: "College of Information Science, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
haunted_places <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-10-10/haunted_places.csv')
#| message: false
```

**Data Discription**

The Haunted Places dataset contains information about haunted locations across the U.S., 
including site names, cities, states, descriptions, and geographic coordinates (latitude and longitude). 
It’s perfect for exploring spatial patterns, visualizing haunted places on maps, and analyzing the distribution of spooky sites.

Dimensions: The dataset contains 10,992 rows and 10 columns. Each row represents a unique haunted site, while the columns capture various details such as descriptions of the hauntings, geographic coordinates (latitude and longitude), and location information (city, state).

**Why This Data Set**
Diverse Data Types: The dataset includes both text (descriptions, locations) and numeric data (coordinates), 
allowing you to practice data cleaning, handling, and analysis across different data types.

Key Variables:

city: The city where the place is located.
state: The U.S. state where the place is located.
location: The title or name of the haunted place.
description: A textual description of the place (varying levels of detail).
longitude and latitude: The geographical coordinates of the haunted location.
city_longitude and city_latitude: The geographical coordinates of the city center.
state_abbrev: Two-letter state abbreviations for easy reference.
The dataset focuses on the United States, with each entry representing a haunted place in various cities and states across the country. This rich combination of geographical and descriptive data allows for an exploration of trends in haunted places, from regional distributions to deeper historical connections.


Creative Visualizations: With its unique theme, the dataset offers opportunities to create visually striking and
thematic data visualizations, such as interactive maps or storytelling through data, making it ideal for engaging projects.

'''{r}
#dimensions of dataset

dim(haunted_places)
'''

'''{r}
#Summary of dataset

summary(haunted_places)
'''
The dataset includes a variety of variables such as

Location: Name or title of the haunted site.
Description: Details of the haunting events.
State: U.S. state where the site is located.
Latitude and Longitude: Geographic coordinates for mapping.
City: Urban or rural location context.


'''{r}
#Data preview

head(haunted_place)
'''


## Questions

1. Geographical Analysis: What are the most haunted states and cities in the United States? Is there a relationship between a state's population and the number of haunted locations?
Variables: state, city, location, longitude, latitude, state_abbrev, and external population data.
Plan: Use the geographical data (e.g., longitude, latitude, state, and city) to plot haunted places across the U.S. A heatmap or bubble chart will visualize areas with the highest density of haunted places. External population data for each state will be merged to check for correlations between population and haunted places using scatter plots and correlation analysis.

2. Historical Significance: Do locations with detailed historical backgrounds in the description tend to be more haunted? Are there certain types of places (e.g., cemeteries, public buildings) that have more frequent hauntings based on the description data?
Variables: description, location, and type (if derived from description).
Plan: Perform text analysis on the description column to classify haunted places based on historical significance or the amount of historical detail provided. We will also explore if certain locations (e.g., older places, battlefields, or public buildings) are more frequently associated with hauntings. Text mining techniques (e.g., word frequency analysis) may help identify historical themes.
These approaches will help in understanding both the geographical spread of haunted places and the historical significance tied to these locations.

## Work Assignment

| Task Name                         | Status      | Assignee        | Due    | Priority  | Summary                                                                |
|-----------------------------------|-------------|-----------------|--------|-----------|------------------------------------------------------------------------|
| Choose Dataset & Define Questions | Completed   | Team            | Week 1 | High      | Complete dataset overview and finalize questions.                      |
| Proposal Draft                    | Completed   | Disha, Sindhu   | Week 2 | High      | Draft proposal with dataset description, questions, and plan.          |
| Revise Proposal                   | Not Started | Team            | Week 3 | High      | Revise proposal based on feedback received.                            |
| Data Analysis for Question 1      | Not Started | Sahana          | Week 4 | Moderate  | Perform analysis and visualizations for Question 1.                    |
| Data Analysis for Question 2      | Not Started | Sindhu          | Week 4 | High      | Perform analysis and visualizations for Question 2.                    |
| Write-Up Draft                    | Not Started | Team            | Week 5 | Moderate  | Draft the final write-up of findings.                                  |
| Presentation Preparation          | Not Started | Team            | Week 6 | High      | Finalize report, prepare presentation materials, and present findings. |

