---
title: "Haunted Places in the United States"
subtitle: "Proposal"
author: 
  - name: "The Data Drafters"
    affiliations:
      - name: "College of Information Science, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
haunted_places <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-10-10/haunted_places.csv')
#| message: false
```

**Data Discription**

The Haunted Places dataset contains information about haunted locations across the U.S., 
including site names, cities, states, descriptions, and geographic coordinates (latitude and longitude). 
It’s perfect for exploring spatial patterns, visualizing haunted places on maps, and analyzing the distribution of spooky sites.

**Why This Data Set**
Diverse Data Types: The dataset includes both text (descriptions, locations) and numeric data (coordinates), 
allowing you to practice data cleaning, handling, and analysis across different data types.

Creative Visualizations: With its unique theme, the dataset offers opportunities to create visually striking and
thematic data visualizations, such as interactive maps or storytelling through data, making it ideal for engaging projects.

'''{r}
#dimensions of dataset

dim(haunted_places)
'''

'''{r}
#Summary of dataset

summary(haunted_places)
'''
The dataset includes a variety of variables such as

Location: Name or title of the haunted site.
Description: Details of the haunting events.
State: U.S. state where the site is located.
Latitude and Longitude: Geographic coordinates for mapping.
City: Urban or rural location context.


'''{r}
#Data preview

head(haunted_place)
'''


## Questions

Question 1

How are haunted places distributed across different U.S. states? 
Are there specific states or regions with a higher concentration of haunted locations?

Why this question

We’re looking at how haunted places are spread out across different states to see if some states have more haunted sites than others. 
This can help us understand if certain regions are more likely to have these sites and why that might be.

Variable to be created

State Density: This variable does not currently exist in the dataset but could be created if 
              additional data on state population or area is obtained.

Question 2
What common themes or types of hauntings are reported in the descriptions? 
Are there patterns or trends in the types of hauntings based on the location or state?

Why this question

We want to see if there are common themes or types of hauntings described in the dataset. 
This can show us if there are recurring patterns or types of supernatural events in different places.

## Analysis Plan

| Task Name                         | Status      | Assignee        | Due    | Priority  | Summary                                                                |
|-----------------------------------|-------------|-----------------|--------|-----------|------------------------------------------------------------------------|
| Choose Dataset & Define Questions | Completed   | Team            | Week 1 | High      | Complete dataset overview and finalize questions.                      |
| Proposal Draft                    | Completed   | Disha, Sindhu   | Week 2 | High      | Draft proposal with dataset description, questions, and plan.          |
| Revise Proposal                   | Not Started | Team            | Week 3 | High      | Revise proposal based on feedback received.                            |
| Data Analysis for Question 1      | Not Started | Sahana          | Week 4 | Moderate  | Perform analysis and visualizations for Question 1.                    |
| Data Analysis for Question 2      | Not Started | Sindhu          | Week 4 | High      | Perform analysis and visualizations for Question 2.                    |
| Write-Up Draft                    | Not Started | Team            | Week 5 | Moderate  | Draft the final write-up of findings.                                  |
| Presentation Preparation          | Not Started | Team            | Week 6 | High      | Finalize report, prepare presentation materials, and present findings. |

